### 2020-11-18 manuscript/preprint_Luken_Kucharsky_Visser_Classifying_Eye_Movement_Events.Rmd
upload preprint

authornote: |
 We would like to thank Daan van Renswoude, Maartje Raijmakers, and Maximilian Maier for their helpful comments on earlier versions of this paper. Furthermore, we would like to acknowledge feedback by Karel Veldkamp and Phil Norberts in the early stage of this project. Šimon Kucharský was supported by the NWO (Nederlandse Organisatie voor Wetenschappelijk Onderzoek) grant no. 406.10.559.

@@

keywords          : "eye-tracking; event detection; parameter recovery; dependent mixture models"
wordcount         : "10327"

@@

@@

```{r plot-algorithm-workflow, fig.cap="(ref:plot-algorithm-workflow)", fig.height = 8}

@@

p2 <- ggplot(toPlot, aes(x = t, y = vel)) + 
  geom_line() + 
  geom_point(size = 2) +
  scale_y_continuous(name = "Velocity (deg/s)", limits = c(0, 250), breaks = pretty(c(0, 250)))

p3 <- ggplot(toPlot, aes(x = t, y = vel)) + 
  geom_line() + 
  geom_point(aes(color = event.pre), size = 2) +
  scale_y_continuous(name = "Velocity (deg/s)", limits = c(0, 250), breaks = pretty(c(0, 250)))

p4 <- ggplot(toPlot, aes(x = t, y = vel)) + 
  geom_line() + 
  geom_point(aes(color = event.post), size = 2) +
  scale_y_continuous(name = "Velocity (deg/s)", limits = c(0, 250), breaks = pretty(c(0, 250)))

@@

@@

sim.pars.trans <- as.data.frame(matrix(c("Interval", "-", "[.01,.99]", "1-a(i=j)/(k-1)",
                                           "Default", "1/k", "0.9", "0.1/(k-1)"), nrow = 2, byrow = T))

apa_table(sim.pars.trans,
          col.names = c("", "Initial\nstate\nprob.", "Trans. prob.\nsame state", "Trans. prob.\nother state"),
          caption = "(ref:sim-parameters-trans)",
          note = "(ref:sim-parameters-trans-note)")

@@

apa_table(cbind(rep(c("Interval", "Default"), 4), sim.pars.resp),
          col.names = c("", "Shape", "Scale", "Shape", "Scale", "Mean", "Concentration"),
          col_spanners = list(Velocity = c(2, 3), Acceleration = c(4, 5), "Rel. angle" = c(6, 7)),
          stub_indents = list("State 1" = c(1, 2), "State 2" = c(3, 4), "State 3" = c(5, 6), "State 4" = c(7, 8)),
          caption = "(ref:sim-parameters-resp)",
          note = "(ref:sim-parameters-resp-note)")

@@

(ref:sim-parameters-trans) Intervals and Default Parameter Values for the Transition Model in the Simulation
(ref:sim-parameters-trans-note) The transition probability for staying in the same state is denoted by $a_{i=j}$ and the probability for switching to a different state by $a_{i\neq j}$. The number of states in the model is denoted by *k*.

@@

(ref:tab-starting-values) Starting Values for the Response Model for Fitting gazeHMM to Benchmark Data
(ref:tab-starting-values-note) Starting values for velocity and acceleration signals are shown before scaling down by factor 100.

@@

colnames(tab.start) <- c("Shape", "Scale", "Shape", "Scale", "Mean", "Concentration")
rownames(tab.start) <- c("Fixation", "Saccade", "PSO", "Pursuit", "Event 5")

apa_table(as.data.frame(tab.start),
          col_spanners = list(Velocity = c(2, 3), Acceleration = c(4, 5), "Rel. angle" = c(6, 7)),
          format.args = list(na_string = "-", drop0trailing = T),
          caption = "(ref:tab-starting-values)",
          note = "(ref:tab-starting-values-note)")

@@

(ref:plot-example-classification) Example data from @Andersson2017 displayed as x-, and y-coordinates, velocity (in deg/s), acceleration (in deg/s$^2$), and sample-to-sample (relative) angle (in rad). Colors indicate classification by gazeHMM and symbols by human coder MN. Note that the human coder classified all fixations by gazeHMM as smooth pursuits and delayed the separation between the saccade and PSO.

```{r plot-example-classification, fig.cap="(ref:plot-example-classification)", fig.height=7.5}

@@

# Declarations
## Funding
Šimon Kucharský was supported by the NWO (Nederlandse Organisatie voor Wetenschappelijk Onderzoek) grant no. 406.10.559.

## Conflicts of Interest
The authors declare that they have no financial or non-financial interests regarding the content of this article.

## Ethics Approval
Not applicable.

## Consent to Participate
Not applicable.

## Consent to Publication
Not applicable.

## Availability of Data and Materials
See Open Practices and Data Availability Statement.

## Code Availability
See Open Practices and Data Availability Statement.

>>>>

authornote: |
 *Please note that this is a preprint and has not been certified by peer review*.

 We would like to thank Daan van Renswoude, Maartje Raijmakers, and Maximilian Maier for their helpful comments on earlier versions of this paper. Furthermore, we would like to acknowledge feedback by Karel Veldkamp and Phil Norberts in the early stage of this project. Šimon Kucharský was supported by the NWO (Nederlandse Organisatie voor Wetenschappelijk Onderzoek) grant no. 406.10.559.

@@

keywords          : "eye-tracking; event detection; parameter recovery; dependent mixture models"
wordcount         : "10279"

@@

\raggedbottom

@@

```{r plot-algorithm-workflow, fig.cap="(ref:plot-algorithm-workflow)", fig.height = 7}

@@

p2 <- ggplot(toPlot, aes(x = t, y = vel)) + 
  geom_line() + 
  geom_point(size = 2) +
  scale_y_continuous(name = "Velocity (deg/s)", limits = c(0, 300), breaks = c(0, 100, 200, 300))

p3 <- ggplot(toPlot, aes(x = t, y = vel)) + 
  geom_line() + 
  geom_point(aes(color = event.pre), size = 2) +
  scale_y_continuous(name = "Velocity (deg/s)", limits = c(0, 300), breaks = c(0, 100, 200, 300))

p4 <- ggplot(toPlot, aes(x = t, y = vel)) + 
  geom_line() + 
  geom_point(aes(color = event.post), size = 2) +
  scale_y_continuous(name = "Velocity (deg/s)", limits = c(0, 300), breaks = c(0, 100, 200, 300))

@@

(ref:sim-parameters-trans) Intervals and Default Parameter Values for the Transition Model in the Simulation
(ref:sim-parameters-trans-note) The initial state probability is denoted by $\rho_i$. The transition probability for staying in the same state is denoted by $a_{i=j}$ and the probability for switching to a different state by $a_{i\neq j}$. The number of states in the model is denoted by *k*.

@@

sim.pars.trans <- as.data.frame(matrix(c("Interval", "-", "[0.01,0.99]", "(1-a(i=j))/(k-1)",
                                           "Default", "1/k", "0.9", "0.1/(k-1)"), nrow = 2, byrow = T))

apa_table(sim.pars.trans,
          align = "c",
          col.names = c("", "$\\rho_i$", "$a_{i=j}$", "$a_{i \\neq j}$"),
          caption = "(ref:sim-parameters-trans)",
          note = "(ref:sim-parameters-trans-note)")

@@

apa_table(cbind(rep(c("Interval", "Default"), 4), sim.pars.resp),
          align = "c",
          col.names = c("", "$\\alpha$", "$\\beta$", "$\\alpha$", "$\\beta$", "$\\mu$", "$\\kappa$"),
          col_spanners = list(Velocity = c(2, 3), Acceleration = c(4, 5), "Rel. angle" = c(6, 7)),
          stub_indents = list("State 1" = c(1, 2), "State 2" = c(3, 4), "State 3" = c(5, 6), "State 4" = c(7, 8)),
          caption = "(ref:sim-parameters-resp)",
          note = "(ref:sim-parameters-resp-note)")

@@

@@

(ref:tab-starting-values) Starting Values for the Response Model for Fitting gazeHMM to Benchmark Data
(ref:tab-starting-values-note) Starting values for velocity and acceleration signals are shown before scaling down by factor 100. Shape parameters are denoted by $\alpha$, scale parameters by $\beta$, mean parameters by $\mu$, and concentration parameters by $\kappa$.

@@

#colnames(tab.start) <- c("$\\alpha$", "$\\beta$", "$\\alpha$", "$\\beta$", "$\\mu$", "$\\kappa$")
rownames(tab.start) <- c("Fixation", "Saccade", "PSO", "Pursuit", "Event 5")

apa_table(as.data.frame(tab.start),
          align = "c",
          col.names = c("", "$\\alpha$", "$\\beta$", "$\\alpha$", "$\\beta$", "$\\mu$", "$\\kappa$"),
          col_spanners = list(Velocity = c(2, 3), Acceleration = c(4, 5), "Rel. angle" = c(6, 7)),
          format.args = list(na_string = "-", drop0trailing = T),
          caption = "(ref:tab-starting-values)",
          note = "(ref:tab-starting-values-note)")

@@

(ref:plot-example-classification) Example data from @Andersson2017 displayed as x-, and y-coordinates (in deg), velocity (in deg/s), acceleration (in deg/s$^2$), and sample-to-sample (relative) angle (in rad). Colors indicate classification by gazeHMM and symbols by human coder MN. Note that the human coder classified all fixations by gazeHMM as smooth pursuits and delayed the separation between the saccade and PSO.

```{r plot-example-classification, fig.cap="(ref:plot-example-classification)", fig.height=7}

@@
